Characterization of gaps in directly bonded Si compound optics

### 2014-03-20 AO_bondedSi.tex
commit before I remove the double beam mode detailed discussion

We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.  The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.  We bored divots in silicon with inductively coupled plasma etching.  The tool was a Unaxis.  We used two different gases with low and high etch rates.  CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.  The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.  We protected the unexposed areas with thick photoresist.  We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.  Figure \ref{figX} shows a schematic of VG03 and its pattern.  In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates.  For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{fig}.  The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.  The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.

@@

Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.

@@

\subsection{The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon}

@@

The Si$-$air$-$Si cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  The computation is simply Fresnel's law at normal incidence:

$$
R = \frac{(n-1)^2}{(n+1)^2} \\
F \equiv \frac{4R}{(1-R)^2}
$$

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.

\subsection{Consideration of multiple incoherent reflections}
\label{secTheory}
Multiple incoherent reflections occur within the Si substrate.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

Consider a pair of bonded Si pucks with a gap thickness $d$. We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.  We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.  We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.

@@

\section{Measurement strategy, conditions, and details}
The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \nocite{cary5000web}.  The typical measurement parameters and tool performance are listed below.  More details about the tool are available at the vendor's website.  Briefly, the tool has a double monochromator.  The double monochromator feature reduces scattered light, with the vendor reported linearity exceeding 40 dB.

>>>>

We evaluated our metrology method by imbedding gaps of known sizes in the interface.  Here we describe the process of how we made the gaps and how we characterized their axial extent.  The three important characteristics of the gaps are their axial extent, their lateral areal extent, and their areal fill factor.  We bored divots in silicon with inductively coupled plasma etching.  The tool was a Unaxis.  We used two different gases with low and high etch rates.  CHF$_3$ on Si exhibited about 0.3 nm/second etch rate, SF$_6$ exhibited about 1 micron/minute.  The area exposed to the process gases was about 1 inch diameter circles, or 15 mm $\times$ 20 mm rectangles.  We protected the unexposed areas with thick photoresist.  We achieved etch depths of 15$\pm$5 nm and 4000$\pm$100 nm, as measured with Veeco NT9100 Optical Profiler for the small depth, and Dektak stylus profilometry for the large depth.  Figure \ref{figX} shows a schematic of VG03 and its pattern.  In a second round of experiments we achieved gap sizes of 14$-$95 nm on four substrates.  For measurements taken with the Veeco, uncertainties were constructed by inspecting the histogram of topology, as shown in Figure \ref{fig}.  The large-scale distortions were removed from the topology by masking and flattening to ensure that the uncertainties accurately reflect the distribution of measured heights.  The stylus profiliometry measured values were simply assigned an uncertainty of 5\%, which was consistent with our intuition and experience.  Table \ref{tabbondexper} lists part numbers and descriptions for our bonded substrates.

@@

@@

\subsection{The air cavity is a low Finesse Fabry-P\`{e}rot etalon}

@@

The air cavity is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  Note that we imprecisely refer to the gap as an air gap, though in principle the gap has an unknown mixture of gases for example water vapor, or other vapors that were outgassed from reactions catalyzed by hydrocarbons.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We assume Si is nonabsorbing longward of $\lambda$ = 1150 nm.  We compute the reflectivity for an Si$-$air interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  The computation is simply Fresnel's law at normal incidence:

\begin{eqnarray}
R = \frac{(n-1)^2}{(n+1)^2} \\
F \equiv \frac{4R}{(1-R)^2}
\end{eqnarray}

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of Finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of Finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.  In the next subsection we work out the wavelength dependent transmission for the air etalon

\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
\label{secTheory}
There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si puck with no air gap, and second a pair of bonded Si pucks with a gap thickness $d$.  The DSP Si puck with no air gap has a transmission equal to:

@@

Now consider a pair of bonded Si pucks with a gap thickness $d$. We derive the absolute transmission bonded Si wafers with a gap in Equation \ref{eqn:Tetalon} in the Appendix.  We normalize the absolute transmission by the transmission with no gap to isolate the effect of the gap.  We call this normalized transmission the etalon transmission $T_{e}$, since it isolates the effect of the etalon on the transmission.

@@

\section{Measurement strategy, conditions, and details}
The measurements were performed with an Agilent Cary 5000 UV-Vis-NIR spectrophotometer \cite{cary5000web}.  The typical measurement parameters and tool performance are listed below.  More details about the tool are available at the vendor's website.  Briefly, the tool has a double monochromator.  The double monochromator feature reduces scattered light.  The vendor reported linearity exceeding 40 dB.
